Structural Insights 2023: A Linear Analysis of Apartment Age and Safety Scores in Toronto\thanks{Code and data supporting this analysis are available at https://github.com/TEJMaster/Toronto-Building-Safety-Analysis-2023.git}

### 2024-03-15 paper/paper.tex
some syntax modification

---
title: "Structural Insights 2023: A Linear Analysis of Apartment Age and Safety Scores in Toronto"
title-block-banner: true
abstract: "This study harnesses a dataset from the Toronto Open Data portal, comprising 3,435 records of apartment evaluations, to analyze the correlation between building age and safety scores. Employing a linear regression model, we discovered a significant inverse relationship, highlighting that older buildings tend to have lower safety evaluations. The findings suggest a need for policy reform, particularly in the maintenance and inspection of Toronto's aging residential infrastructure. Despite the study’s focus on building age, further research considering additional variables could offer a more comprehensive understanding of factors influencing building safety."
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Toronto-Building-Safety-Analysis-2023.git"
author: "Terry Tu, Jingyi Shen, Yaning Jin "
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
```

# Introduction

The structural integrity of urban buildings is a paramount concern for city planners, policymakers, and residents alike. The tragic collapse of a condominium in South Florida in June 2021, which resulted in 98 casualties, serves as a stark reminder of the potential consequences of neglecting building maintenance and inspections [@Anderson_2021]. This event highlights a critical gap in our understanding of the factors that contribute to the safety and longevity of urban residential structures. In light of this, our study aims to investigate the relationship between building age and evaluation scores in the context of Toronto's residential buildings.

@@

The remainder of this paper is structured as follows: @sec-data (Data) details the raw data, the cleaning process, and provides an overview of the data distribution. @sec-model (Model) describes the linear model employed to predict the building evaluation score based on the building's age. @sec-result (Results) presents the coefficients of the linear model, analyses the residuals, and includes a plot illustrating the linear model applied to the analysis dataset. Finally, @sec-discussion (Discussion) discusses the limitations of the analysis and suggests potential insights for future research.

# Data {#sec-data}

```{r include=FALSE, message=FALSE}
# read the data file and store it in the 'data' variable
cleaned_data <- read_csv(here("data", "analysis_data", "cleaned_building_data.csv"))
```

## Raw Data

In this research, we examine apartment data sourced using the `opendatatoronto` package [@Toronto_2023]. The dataset encompasses 3,435 records, focusing on specific variables: the ward of each apartment, its evaluation score, and the construction year. The data covers evaluations across Toronto's 25 wards, with scores varying between a minimum of 0 and some achieving the maximum score of 100. The study includes apartments that are over three stories in height and contain at least 10 units. These apartments span a wide range of construction years, from as early as 1805 to as recent as 2023. All the apartments in this dataset adhere to the City's Apartment Building Standards bylaw, which is designed to safeguard the interests and safety of both landlords and tenants by mandating these evaluations.

## Data Analysis Tools

The data analysis was performed using R [@r], a powerful open-source statistical programming language. A suite of packages from the tidyverse [@rTidyverse], an assemblage of R packages designed for data science, was harnessed to enhance the efficiency of our data operations. The `ggplot2` package [@rGgplot2] facilitated the creation of sophisticated visualizations, while `dplyr` [@rDplyr] provided a grammar of data manipulation, offering a coherent set of verbs that help in filtering, summarizing, and arranging the dataset. The `readr` package [@rReadr] was utilized for its fast and friendly data reading capabilities. Navigation and file path management were streamlined using the `here` package [@R-here], simplifying the process of file referencing within the project's directory structure. Report generation was dynamically handled by knitr [@rKnitr], enabling the integration of R code within this document. Additionally, `kableExtra`[@R-kableExtra] was employed to produce aesthetically pleasing and customizable tables, enriching the presentation of our results. For the Bayesian analysis, rstanarm [@R-rstanarm] was utilized to create a linear model, providing an elegant interface to Stan, a state-of-the-art platform for statistical modeling and high-performance statistical computation. This package enabled us to estimate the relationship between building age and safety evaluation scores using a Bayesian framework.

## Variable Description

@@

## Sample of Cleaned Building Evaluation Data

```{r}
#| label: tbl-data-sample
#| tbl-cap: Sample of Building Evaluation Data
#| echo: false

sample_data <- head(cleaned_data) %>%
  select(X_id, AGE, CURRENT.BUILDING.EVAL.SCORE)

# Create a nicely formatted table with custom column names
kable(sample_data, col.names = c("Building ID", "Age (Years)", "Evaluation Score"))
```
@tbl-data-sample represents a subset of the larger building evaluation data. Each row corresponds to a unique building identified by X_id. The AGE column indicates the age of the building in years, while the CURRENT.BUILDING.EVAL.SCORE column shows the building's current evaluation score on a scale from 0 to 100, with higher scores indicating better compliance with safety standards.

## Measurement:

@@

## Data Exploration:

### Data Summary
```{r}
#| label: tbl-summary-table
#| tbl-cap: Summary Statistics for the Cleaned Dataset
#| echo: false

summary_data <- cleaned_data %>%
  summarise(
    Count = n(),
    Mean_Age = mean(AGE, na.rm = TRUE),
    Std_Age = sd(AGE, na.rm = TRUE),
    Mean_Score = mean(CURRENT.BUILDING.EVAL.SCORE, na.rm = TRUE),
    Std_Score = sd(CURRENT.BUILDING.EVAL.SCORE, na.rm = TRUE)
  ) %>%
  mutate(across(everything(), ~round(., 3)))

# Create and print a nicely formatted table
summary_table <- kable(summary_data, 
                       col.names = c("Count", "Mean Age (Years)", "SD of Age (Years)", 
                                     "Mean Score", "SD of Score"), 
                       align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                position = "center") %>%
  column_spec(1, bold = TRUE) # Make the first column bold

# Print the summary table
summary_table
```

@tbl-summary-table presents a concise overview of the key descriptive metrics for the age of buildings and their evaluation scores within the dataset:

- **Count**: The dataset encompasses a total of 1,748 buildings, indicating a robust sample size for statistical analysis.

- **Mean Age**: The average age of the buildings is approximately 60.644 years, suggesting that the dataset primarily includes buildings that have been standing for over half a century.

- **Standard Deviation of Age**: With a standard deviation of about 19.939 years, there is considerable variability in the ages of the buildings. This range implies a diverse sample, encompassing both relatively new constructions and much older structures.

- **Mean Score**: The average building evaluation score is approximately 87.504 on a scale from 0 to 100. A higher score corresponds to better adherence to safety and maintenance standards, suggesting that the buildings in this dataset are generally well-maintained.

- **Standard Deviation of Score**: The standard deviation of approximately 8.693 for the evaluation scores indicates that while scores are clustered around a high mean, there is still notable variability, reflecting differences in the condition or compliance of the buildings.

@@

### Age Distribution
```{r fig-age-distribution, echo=FALSE, fig.cap="Distribution of building's age for the cleaned data set"}
# Histogram for the distribution of building age
ggplot(cleaned_data, aes(x = AGE)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  labs(
    x = "Age of Building (Years)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

@fig-age-distribution illustrates the frequency distribution of the ages of buildings within the dataset. The x-axis represents the age of the buildings in years, while the y-axis shows the frequency of buildings within each age range. The distribution appears to be unimodal with a peak around the 50-year mark, indicating that a substantial number of the buildings within this dataset were constructed around half a century ago. Following this peak, the frequency gradually decreases, showing that fewer buildings have been in existence beyond this age. The distribution suggests a relatively young building population, with ages concentrated below 100 years, and very few buildings reaching beyond 120 years of age. This visualization aids in understanding the age profile of buildings under consideration for safety evaluation scoring.

### Score Distribution

```{r fig-score-distribution, echo=FALSE, fig.cap="Distribution of building's evaluation score for the cleaned data set"}
# Histogram for CURRENT.BUILDING.EVAL.SCORE distribution
ggplot(cleaned_data, aes(x = CURRENT.BUILDING.EVAL.SCORE)) +
  geom_histogram(binwidth = 5, fill = "darkgreen", color = "white") +
  labs(
    x = "Current Building Evaluation Score",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )
```

@fig-score-distribution showcases the distribution of building evaluation scores. The x-axis represents the scores assigned to buildings, which can range from 0 to 100, with the y-axis displaying the frequency of buildings for each score interval. The chart reveals a concentration of high evaluation scores, suggesting that many of the buildings in the dataset are rated as being in good or excellent condition according to the evaluation criteria. The tallest bars are clustered towards the higher end of the score range, indicating that buildings with scores close to 100 are prevalent. This positive skew in the distribution might reflect effective building standards and maintenance practices in place. The spread of the scores and their frequency can offer insights into the overall health of the evaluated buildings, influencing urban planning and policy development.

# Model {#sec-model}

@@

## Model set-up

Let $y_i$ represent the safety evaluation score for each building. The building's age is denoted by $x_i$.The model is defined as follows:

@@

The model is implemented in R using the rstanarm package, which allows us to incorporate prior beliefs about the parameters and to estimate the posterior distributions. For the intercept $\beta_0$ and slope $\beta_1$ , we choose normal prior distributions centered around our initial estimates based on exploratory data analysis. The `stan_glm` function from `rstanarm` is utilized to carry out the Bayesian regression, providing a framework for inference that incorporates both the data and our prior knowledge.

The priors are chosen to reflect reasonable beliefs about the parameters before observing the data. The intercept prior reflects an expectation that a new building will have a high safety score around 90, with a wide spread to accommodate uncertainty. Similarly, the slope prior indicates an expected small negative impact of age on the safety score, acknowledging some uncertainty in the precise effect. The `stan_glm` function combines these priors with the data to estimate the posterior distributions of the model parameters.

## Model justification

We hypothesize that the coefficient for building age ($\beta_1$) will be negative, reflecting the tendency for older buildings to exhibit lower safety scores. This expectation is rooted in the general understanding that buildings can deteriorate over time due to wear, aging materials, and possible obsolescence in older construction methods. The intercept ($\beta_0$), representing the baseline evaluation score for a new building, is presumed to be relatively high, indicating the score a building would receive if it were built in the current period with modern standards. The slope ($\beta_1$) is therefore of particular interest, as it encapsulates the average annual decline in evaluation score attributable to aging. The Bayesian framework allows us to incorporate these hypotheses as priors in our model, thereby aligning the analysis with our substantive expectations and providing a more nuanced interpretation of the data.

# Result {#sec-result}

```{r echo=FALSE}
# Retrieve the Bayesian linear model from the saved .rds file
bayesian_model <- readRDS(file = here("models", "bayesian_linear_model.rds"))
```

## Model Coefficients Interpretation

```{r}
#| label: tbl-coefficient-table
#| tbl-cap: Summary Statistics for the Coefficients of the Linear Model
#| echo: false

# Extract coefficients from the Bayesian model
coefficients <- coef(bayesian_model)

# Convert coefficients to a data frame for better formatting
coefficients_df <- data.frame(
  Term = c("Intercept", "Slope"),
  Estimate = unname(coefficients) 
)

# Round the estimates to three digits
coefficients_df$Estimate <- round(coefficients_df$Estimate, 3)

# Create a nicely formatted table
coefficient_table <- kable(coefficients_df,
                           col.names = c("Term", "Estimate"),
                           align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE,
                position = "center")

# Print the summary table
coefficient_table
```

@tbl-coefficient-table provides two key coefficients to understand the relationship between building age and the evaluation score:

**Intercept** ($\beta_0$):  
The model's intercept, $\beta_0$, has an estimate of 93.364. This represents the baseline evaluation score for a hypothetical newly constructed building. The intercept suggests that if a building were constructed in the current year, its expected evaluation score would be around 93.364, assuming all other factors remain constant.

**Slope** ($\beta_1$):  
The slope parameter, $\beta_1$, has an estimate of -0.097. This indicates that, on average, the evaluation score is expected to decrease by approximately 0.097 points for each additional year of a building's age. This negative value suggests a trend where older buildings tend to have lower evaluation scores compared to newer ones.

**Interpretation**:
The analysis suggests a negative relationship between building age and evaluation score. The estimates indicate that newer buildings tend to have higher evaluation scores, which decrease as buildings age. This trend highlights the importance of age as a factor in building evaluations and underscores the need for ongoing maintenance and updates to preserve building safety and integrity over time.

## Model Equation

@@

In this model, $y_i$ represents the predicted evaluation score for the $i$-th building, $x_i$ represents the age for the $i$-th building, $\mu_i$ is the mean of the normal distribution for the $i$-th observation, $\beta_0$ is the intercept, and $\beta_1$ is the slope representing the change in evaluation score per year of building age.

## Analysis of Residuals

```{r}
#| label: tbl-residuals-summary
#| tbl-cap: Summary Statistics for the Residuals of the Bayesian Linear Model
#| echo: false

# Extract residuals from the Bayesian model
residuals <- residuals(bayesian_model)

# Create a summary table for residuals and round to three digits
residuals_rounded <- round(summary(residuals), 3)

# Create a data frame for better formatting
residuals_summary_df <- data.frame(
  Statistic = c("Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max"),
  Value = as.numeric(residuals_rounded)  # Convert to numeric to ensure proper formatting
)

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Structural Insights 2023: A Linear Analysis of Apartment Age and Safety Scores in Toronto},
  pdfauthor={Terry Tu, Jingyi Shen, Yaning Jin},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Structural Insights 2023: A Linear Analysis of Apartment Age and
Safety Scores in Toronto\thanks{Code and data supporting this analysis
are available at
https://github.com/TEJMaster/Toronto-Building-Safety-Analysis-2023.git}}
\author{Terry Tu, Jingyi Shen, Yaning Jin}
\date{March 15, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study harnesses a dataset from the Toronto Open Data portal,
comprising 3,435 records of apartment evaluations, to analyze the
correlation between building age and safety scores. Employing a linear
regression model, we discovered a significant inverse relationship,
highlighting that older buildings tend to have lower safety evaluations.
The findings suggest a need for policy reform, particularly in the
maintenance and inspection of Toronto's aging residential
infrastructure. Despite the study's focus on building age, further
research considering additional variables could offer a more
comprehensive understanding of factors influencing building safety.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, breakable, frame hidden, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The structural integrity of urban buildings is a paramount concern for
city planners, policymakers, and residents alike. The tragic collapse of
a condominium in South Florida in June 2021, which resulted in 98
casualties, serves as a stark reminder of the potential consequences of
neglecting building maintenance and inspections
(\protect\hyperlink{ref-Anderson_2021}{Anderson 2021}). This event
highlights a critical gap in our understanding of the factors that
contribute to the safety and longevity of urban residential structures.
In light of this, our study aims to investigate the relationship between
building age and evaluation scores in the context of Toronto's
residential buildings.

@@

The remainder of this paper is structured as follows:
Section~\ref{sec-data} (Data) details the raw data, the cleaning
process, and provides an overview of the data distribution.
Section~\ref{sec-model} (Model) describes the linear model employed to
predict the building evaluation score based on the building's age.
Section~\ref{sec-result} (Results) presents the coefficients of the
linear model, analyses the residuals, and includes a plot illustrating
the linear model applied to the analysis dataset. Finally,
Section~\ref{sec-discussion} (Discussion) discusses the limitations of
the analysis and suggests potential insights for future research.

\hypertarget{sec-data}{%
\section{Data}\label{sec-data}}

\hypertarget{raw-data}{%
\subsection{Raw Data}\label{raw-data}}

In this research, we examine apartment data sourced using the
\texttt{opendatatoronto} package
(\protect\hyperlink{ref-Toronto_2023}{Toronto 2023}). The dataset
encompasses 3,435 records, focusing on specific variables: the ward of
each apartment, its evaluation score, and the construction year. The
data covers evaluations across Toronto's 25 wards, with scores varying
between a minimum of 0 and some achieving the maximum score of 100. The
study includes apartments that are over three stories in height and
contain at least 10 units. These apartments span a wide range of
construction years, from as early as 1805 to as recent as 2023. All the
apartments in this dataset adhere to the City's Apartment Building
Standards bylaw, which is designed to safeguard the interests and safety
of both landlords and tenants by mandating these evaluations.

\hypertarget{data-analysis-tools}{%
\subsection{Data Analysis Tools}\label{data-analysis-tools}}

The data analysis was performed using R (\protect\hyperlink{ref-r}{R
Core Team 2022}), a powerful open-source statistical programming
language. A suite of packages from the tidyverse
(\protect\hyperlink{ref-rTidyverse}{Wickham et al. 2019}), an assemblage
of R packages designed for data science, was harnessed to enhance the
efficiency of our data operations. The \texttt{ggplot2} package
(\protect\hyperlink{ref-rGgplot2}{Wickham 2016}) facilitated the
creation of sophisticated visualizations, while \texttt{dplyr}
(\protect\hyperlink{ref-rDplyr}{Wickham et al. 2022}) provided a grammar
of data manipulation, offering a coherent set of verbs that help in
filtering, summarizing, and arranging the dataset. The \texttt{readr}
package (\protect\hyperlink{ref-rReadr}{Wickham, Hester, and Bryan
2022}) was utilized for its fast and friendly data reading capabilities.
Navigation and file path management were streamlined using the
\texttt{here} package (\protect\hyperlink{ref-R-here}{Müller 2020}),
simplifying the process of file referencing within the project's
directory structure. Report generation was dynamically handled by knitr
(\protect\hyperlink{ref-rKnitr}{Xie 2014}), enabling the integration of
R code within this document. Additionally,
\texttt{kableExtra}(\protect\hyperlink{ref-R-kableExtra}{Zhu 2021}) was
employed to produce aesthetically pleasing and customizable tables,
enriching the presentation of our results. For the Bayesian analysis,
rstanarm (\protect\hyperlink{ref-R-rstanarm}{Goodrich et al. 2020}) was
utilized to create a linear model, providing an elegant interface to
Stan, a state-of-the-art platform for statistical modeling and
high-performance statistical computation. This package enabled us to
estimate the relationship between building age and safety evaluation
scores using a Bayesian framework.

\hypertarget{variable-description}{%
\subsection{Variable Description}\label{variable-description}}

@@

\hypertarget{sample-of-cleaned-building-evaluation-data}{%
\subsection{Sample of Cleaned Building Evaluation
Data}\label{sample-of-cleaned-building-evaluation-data}}

\hypertarget{tbl-data-sample}{}
\begin{longtable}[]{@{}rrr@{}}
\caption{\label{tbl-data-sample}Sample of Building Evaluation
Data}\tabularnewline
\toprule\noalign{}
Building ID & Age (Years) & Evaluation Score \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Building ID & Age (Years) & Evaluation Score \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 62 & 86 \\
2 & 44 & 94 \\
3 & 54 & 78 \\
4 & 73 & 89 \\
5 & 54 & 99 \\
6 & 53 & 99 \\
\end{longtable}

Table~\ref{tbl-data-sample} represents a subset of the larger building
evaluation data. Each row corresponds to a unique building identified by
X\_id. The AGE column indicates the age of the building in years, while
the CURRENT.BUILDING.EVAL.SCORE column shows the building's current
evaluation score on a scale from 0 to 100, with higher scores indicating
better compliance with safety standards.

\hypertarget{measurement}{%
\subsection{Measurement:}\label{measurement}}

@@

\hypertarget{data-exploration}{%
\subsection{Data Exploration:}\label{data-exploration}}

\hypertarget{data-summary}{%
\subsubsection{Data Summary}\label{data-summary}}

\hypertarget{tbl-summary-table}{}
\begin{longtable}[t]{>{}ccccc}
\caption{\label{tbl-summary-table}Summary Statistics for the Cleaned Dataset }\tabularnewline

\toprule
Count & Mean Age (Years) & SD of Age (Years) & Mean Score & SD of Score\\
\midrule
\textbf{1748} & 60.644 & 19.939 & 87.504 & 8.693\\
\bottomrule
\end{longtable}

Table~\ref{tbl-summary-table} presents a concise overview of the key
descriptive metrics for the age of buildings and their evaluation scores
within the dataset:

\begin{itemize}
\item
  \textbf{Count}: The dataset encompasses a total of 1,748 buildings,
  indicating a robust sample size for statistical analysis.
\item
  \textbf{Mean Age}: The average age of the buildings is approximately
  60.644 years, suggesting that the dataset primarily includes buildings
  that have been standing for over half a century.
\item
  \textbf{Standard Deviation of Age}: With a standard deviation of about
  19.939 years, there is considerable variability in the ages of the
  buildings. This range implies a diverse sample, encompassing both
  relatively new constructions and much older structures.
\item
  \textbf{Mean Score}: The average building evaluation score is
  approximately 87.504 on a scale from 0 to 100. A higher score
  corresponds to better adherence to safety and maintenance standards,
  suggesting that the buildings in this dataset are generally
  well-maintained.
\item
  \textbf{Standard Deviation of Score}: The standard deviation of
  approximately 8.693 for the evaluation scores indicates that while
  scores are clustered around a high mean, there is still notable
  variability, reflecting differences in the condition or compliance of
  the buildings.
\end{itemize}

@@

\hypertarget{age-distribution}{%
\subsubsection{Age Distribution}\label{age-distribution}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-age-distribution-1.pdf}

}

\caption{\label{fig-age-distribution}Distribution of building's age for
the cleaned data set}

\end{figure}

Figure~\ref{fig-age-distribution} illustrates the frequency distribution
of the ages of buildings within the dataset. The x-axis represents the
age of the buildings in years, while the y-axis shows the frequency of
buildings within each age range. The distribution appears to be unimodal
with a peak around the 50-year mark, indicating that a substantial
number of the buildings within this dataset were constructed around half
a century ago. Following this peak, the frequency gradually decreases,
showing that fewer buildings have been in existence beyond this age. The
distribution suggests a relatively young building population, with ages
concentrated below 100 years, and very few buildings reaching beyond 120
years of age. This visualization aids in understanding the age profile
of buildings under consideration for safety evaluation scoring.

\hypertarget{score-distribution}{%
\subsubsection{Score Distribution}\label{score-distribution}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-score-distribution-1.pdf}

}

\caption{\label{fig-score-distribution}Distribution of building's
evaluation score for the cleaned data set}

\end{figure}

Figure~\ref{fig-score-distribution} showcases the distribution of
building evaluation scores. The x-axis represents the scores assigned to
buildings, which can range from 0 to 100, with the y-axis displaying the
frequency of buildings for each score interval. The chart reveals a
concentration of high evaluation scores, suggesting that many of the
buildings in the dataset are rated as being in good or excellent
condition according to the evaluation criteria. The tallest bars are
clustered towards the higher end of the score range, indicating that
buildings with scores close to 100 are prevalent. This positive skew in
the distribution might reflect effective building standards and
maintenance practices in place. The spread of the scores and their
frequency can offer insights into the overall health of the evaluated
buildings, influencing urban planning and policy development.

\hypertarget{sec-model}{%
\section{Model}\label{sec-model}}

@@

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

Let \(y_i\) represent the safety evaluation score for each building. The
building's age is denoted by \(x_i\).The model is defined as follows:

@@

The model is implemented in R using the rstanarm package, which allows
us to incorporate prior beliefs about the parameters and to estimate the
posterior distributions. For the intercept \(\beta_0\) and slope
\(\beta_1\) , we choose normal prior distributions centered around our
initial estimates based on exploratory data analysis. The
\texttt{stan\_glm} function from \texttt{rstanarm} is utilized to carry
out the Bayesian regression, providing a framework for inference that
incorporates both the data and our prior knowledge.

The priors are chosen to reflect reasonable beliefs about the parameters
before observing the data. The intercept prior reflects an expectation
that a new building will have a high safety score around 90, with a wide
spread to accommodate uncertainty. Similarly, the slope prior indicates
an expected small negative impact of age on the safety score,
acknowledging some uncertainty in the precise effect. The
\texttt{stan\_glm} function combines these priors with the data to
estimate the posterior distributions of the model parameters.

\hypertarget{model-justification}{%
\subsection{Model justification}\label{model-justification}}

We hypothesize that the coefficient for building age (\(\beta_1\)) will
be negative, reflecting the tendency for older buildings to exhibit
lower safety scores. This expectation is rooted in the general
understanding that buildings can deteriorate over time due to wear,
aging materials, and possible obsolescence in older construction
methods. The intercept (\(\beta_0\)), representing the baseline
evaluation score for a new building, is presumed to be relatively high,
indicating the score a building would receive if it were built in the
current period with modern standards. The slope (\(\beta_1\)) is
therefore of particular interest, as it encapsulates the average annual
decline in evaluation score attributable to aging. The Bayesian
framework allows us to incorporate these hypotheses as priors in our
model, thereby aligning the analysis with our substantive expectations
and providing a more nuanced interpretation of the data.

\hypertarget{sec-result}{%
\section{Result}\label{sec-result}}

\hypertarget{model-coefficients-interpretation}{%
\subsection{Model Coefficients
Interpretation}\label{model-coefficients-interpretation}}

\hypertarget{tbl-coefficient-table}{}
\begin{longtable}[t]{cc}
\caption{\label{tbl-coefficient-table}Summary Statistics for the Coefficients of the Linear Model }\tabularnewline

\toprule
Term & Estimate\\
\midrule
Intercept & 93.364\\
Slope & -0.097\\
\bottomrule
\end{longtable}

Table~\ref{tbl-coefficient-table} provides two key coefficients to
understand the relationship between building age and the evaluation
score:

\textbf{Intercept} (\(\beta_0\)):\\
The model's intercept, \(\beta_0\), has an estimate of 93.364. This
represents the baseline evaluation score for a hypothetical newly
constructed building. The intercept suggests that if a building were
constructed in the current year, its expected evaluation score would be
around 93.364, assuming all other factors remain constant.

\textbf{Slope} (\(\beta_1\)):\\
The slope parameter, \(\beta_1\), has an estimate of -0.097. This
indicates that, on average, the evaluation score is expected to decrease
by approximately 0.097 points for each additional year of a building's
age. This negative value suggests a trend where older buildings tend to
have lower evaluation scores compared to newer ones.

\textbf{Interpretation}: The analysis suggests a negative relationship
between building age and evaluation score. The estimates indicate that
newer buildings tend to have higher evaluation scores, which decrease as
buildings age. This trend highlights the importance of age as a factor
in building evaluations and underscores the need for ongoing maintenance
and updates to preserve building safety and integrity over time.

\hypertarget{model-equation}{%
\subsection{Model Equation}\label{model-equation}}

@@

In this model, \(y_i\) represents the predicted evaluation score for the
\(i\)-th building, \(x_i\) represents the age for the \(i\)-th building,
\(\mu_i\) is the mean of the normal distribution for the \(i\)-th
observation, \(\beta_0\) is the intercept, and \(\beta_1\) is the slope
representing the change in evaluation score per year of building age.

\hypertarget{analysis-of-residuals}{%
\subsection{Analysis of Residuals}\label{analysis-of-residuals}}
